\large{\bf Quantitative genetics with stable laws }

### 2023-06-11 writeup.tex
start at simulation whatnot

\subsection{Stabilizing Selection}

\subsubsection{Gaussian Solution}

Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
\[
	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
\]
Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)}]$  to conclude that
\[
	e^{i m \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
	= e^{i \frac{m}{1-2 a \sigma^{2}}\theta 
		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
\]
which, equating real and complex components, has a solution provided $m = 0$ and
\[
	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
	=  2\eta^{2}(1-4a\eta^{2}) + O(a^{2}).
\]

\subsubsection{Cauchy Solution}

In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).

If we now suppose that $Z$ is a $\text{Cauchy}(0,\gamma)$ random variable.  Then $Z$ has probability density function $\frac{\gamma}{\pi} \frac{1}{\gamma^{2} + x^{2}}$, whereas 
\[
	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}
	= e^{-\delta |\theta|} = e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|}
\]
which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$.

@@

>>>>

\subsection{Stabilizing Selection}
    \label{sec:stabilizing_selection}

\subsubsection{Gaussian Solution}
    \label{sec:stabilizing_gaussian}

Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
\[
	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
\]
Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  to conclude that
if $\xi$ is Gaussian with mean zero and variance $\eta^2$,
then equation~\eqref{CHARACTERISTIC} is
\comment{something not right here? need a factor of 2 on the left?}
\[
	e^{i m \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
	= e^{i \frac{m}{1-2 a \sigma^{2}}\theta 
		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
\]
which, equating real and complex components, has a solution provided $m = 0$ and
\[
	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
	=  2\eta^{2}(1-4a\eta^{2}) + O(a^{2}).
\]

\subsubsection{Cauchy Solution}
    \label{sec:stabilizing_cauchy}

In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise $\xi$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).

If we now suppose that $Z$ is a $\text{Cauchy}(0,\gamma)$ random variable.  Then $Z$ has probability density function $\frac{\gamma}{\pi} \frac{1}{\gamma^{2} + x^{2}}$, whereas 
\[
	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}
	= e^{-\delta |\theta|} = e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|}
\]
which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$,
and so solves equation~\eqref{CHARACTERISTIC}.

@@

In Section~\ref{sec:stabilizing_selection}
we have found two explicit conditions in which
if the form of segregation noise and stabilizing selection
satisfy certain conditions,
then the limiting phenotype distribution
will be either Gaussian (Subsection~\ref{sec:stabilizing_gaussian}
or Cauchy (Subsection~\ref{sec:stabilizing_cauchy}.
Furthermore, we expect based on
\citet{barton2017infinitesimal} and arguments above
that the distribution of segreation noise
is determined by the tail exponent of the distribution of effect sizes
(under appropriate assumptions; e.g., that most mutations are of small effect).
However, it is not clear how selection would affect these arguments,
especially in the stable case where some number of mutations are expected to
\emph{not} have small effect (and therefore such mutations will not be nearly neutral).

Therefore,
to study how this works in practice, we turn to simulation.
We used SLiM v3 \comment{(cite)}
to simulate a discrete-time approximation to the continuous-time Moran model described above
as follows.
Individuals are diploid and sexual;
the genome is of lenth $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
mutations occur at a rate of $m$ per bp.
Each new mutation is independently assigned an ``effect''
drawn from the ``effect size distribution'' (and will be either Gaussian or Cauchy).
The phenotype of each individual is determined additively
by summing the effects of all mutations they carry across both chromosomes.
Let $f(z) = 1 + \mu(z)$
be the ``fitness'' of an individual with phenotype $z$,
and let $dt$ be a small constant (we took $dt=0.01$).
Then, at each time step
each individual dies with probability $1 - \exp(-dt f(z))$;
if there are $k$ such individuals chosen to die in a given time step,
then we choose $k$ individuals uniformly and without replacement
to produce new offspring;
for each such reproduction event the parent chooses a mate uniformly at random from the population.
This maintains the population at a fixed size, $N$.
(Note that individuals may self and individuals chosen to die may also reproduce,
but in a large population these details should be unimportant.)

We compared two situations:
(1) the effect size distribution was Gaussian with mean zero and standard deviation $\sigma_g$,
and $f(z) = e^{\alpha z^2}$; and
(2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$.
\comment{TODO: just refer to equations above?}
We ran simulations of each with $N=XXX$ for XXX time steps of ``burn-in''
(judged to be long enough as the number of segregating mutations had plateaued),
following which we recorded the phenotypes of each new offspring and their two parents
for XXX time steps (a total of around YYY offspring).

Figure XXX shows the resulting distributions of segregation noise:
i.e., the distribution of child phenotype minus average of parents' phenotypes,
across all offspring following burn-in.
